Types from data: \textnormal{Making structured data first-class citizens in F\#}

### 2016-05-10 pldi-2016/paper.tex
Minor tweaks for arXiv submission

\documentclass[10pt]{sigplanconf}

\usepackage{listings}
\usepackage{enumitem}
\usepackage[hyphens]{url}
\usepackage[svgnames]{xcolor}
\definecolor{lcl}{RGB}{140,0,100}
\usepackage[colorlinks=true,breaklinks,draft=false]{hyperref}
\hypersetup{urlcolor=lcl,linkcolor=lcl,citecolor=lcl}
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}
\usepackage{amssymb}
\usepackage{multicol}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage[fleqn]{amsmath}
\usepackage{graphics}
\usepackage{stmaryrd}
\usepackage{amsthm}
\usepackage{ftnright}
\usepackage[T1]{fontenc}
\usepackage{semantic}
\usepackage{enumitem}
\usepackage[hang,flushmargin]{footmisc}
\usepackage[notquote]{hanging}
\usepackage{flushend}
\usepackage{enumerate}
\usepackage{etoolbox}

@@

\noindent
\begin{equation*}
\begin{array}{l}
 \kvd{let}~\ident{doc}=\ident{Http.Request}(\str{http://api.owm.org/?q=NYC}) \\
 \kvd{match}~\ident{JsonValue.Parse}(\ident{doc})~\kvd{with} \\
 |~\ident{Record}(\ident{root})\rightarrow \\
 \quad \kvd{match}~\ident{Map.find}~\str{main}~\ident{root}~\kvd{with} \\
 \quad |~\ident{Record}(\ident{main})\rightarrow \\
 \quad \quad \kvd{match}~\ident{Map.find}~\str{temp}~\ident{main}~\kvd{with} \\
 \quad \quad |~\ident{Number}(\ident{num})\rightarrow \ident{printfn}~\str{Lovely \%f!}~\ident{num} \\
 \quad \quad |~\_\rightarrow \ident{failwith}~\str{Incorrect format} \\
 \quad |~\_\rightarrow \ident{failwith}~\str{Incorrect format} \\
 |~\_\rightarrow \ident{failwith}~\str{Incorrect format}
\end{array}
\end{equation*}
%
The code assumes that the response has a particular shape described in the documentation. The
root node must be a record with a \strf{main} field, which has to be another record containing
a numerical \strf{temp} field representing the current temperature. When the shape is different,
the code fails. While not immediately unsound, the code is prone to errors if strings are
misspelled or incorrect shape assumed.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.80,trim=5mm 5mm 5mm 5mm,clip]{images/hierarchy.pdf} % left bottom right top
\end{center}
\vspace{-0.5em}
\caption{Important aspects of the preferred shape relation}
\label{fig:subtyping-diagram}
\end{figure}

@@

Such $e'$ is obtained by transforming sub-expressions of $e$ using one of the following
translation rules:
%
\begin{enumerate}[(i)]
\item
$C[e]$ to $C[\kvd{match}~e~\kvd{with}~\ident{Some}(v) \rightarrow v~|~\ident{None} \rightarrow \ident{exn}]$
\item
$C[e]$ to $C[e.\ident{M}]$ where $\ident{M} = \tytagof(\sigma)$ for some $\sigma$
\item
$C[e]$ to $C[\ident{int}(e)]$
\end{enumerate}
\end{remark}

@@

\begin{itemize}
\item[--] Primitive shapes can become nullable (i), \ident{int} can become \ident{float} (iii)
  or become a part of a labelled top type (ii).
\item[--] Record shape fields can change shape (recursively) and record may become a part
  of a labelled top type (ii).
\item[--] For list and nullable shapes, the shape of the value may change (we apply the
  transformations recursively).
\item[--] For the $\kvd{any}$ shape, the original code will continue to work (none of the labels is ever removed).
\end{itemize}
\vspace{-1.5em}
\end{proof}
%
\noindent
Intuitively, the first transformation is needed when the new sample makes a type optional.
This happens when it contains a \kvd{null} value or a record that does not contain a field
that all previous samples have. The second transformation is needed when a shape $\sigma$
becomes $\kvd{any}\langle\sigma, \ldots\rangle$ and the third one is needed when $\ident{int}$
becomes \ident{float}.

>>>>

\documentclass[10pt,nocopyrightspace]{sigplanconf}
\pdfoutput=1
\usepackage{listings}
\usepackage{enumitem}
\usepackage[hyphens]{url}
\usepackage[svgnames]{xcolor}
\definecolor{lcl}{RGB}{140,0,100}
\usepackage[colorlinks=true,breaklinks,draft=false]{hyperref}
\hypersetup{urlcolor=lcl,linkcolor=lcl,citecolor=lcl}
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}
\usepackage{amssymb}
\usepackage{multicol}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage[fleqn]{amsmath}
\usepackage{graphics}
\usepackage{stmaryrd}
\usepackage{amsthm}
\usepackage{ftnright}
\usepackage[T1]{fontenc}
\usepackage{semantic}
\usepackage{enumitem}
\usepackage[hang,flushmargin]{footmisc}
\usepackage[notquote]{hanging}
\usepackage{flushend}
\usepackage{etoolbox}

@@

\begin{equation*}
\begin{array}{l}
 \kvd{let}~\ident{doc}=\ident{Http.Request}(\str{http://api.owm.org/?q=NYC}) \\
 \kvd{match}~\ident{JsonValue.Parse}(\ident{doc})~\kvd{with} \\
 |~\ident{Record}(\ident{root})\rightarrow \\
 \quad \kvd{match}~\ident{Map.find}~\str{main}~\ident{root}~\kvd{with} \\
 \quad |~\ident{Record}(\ident{main})\rightarrow \\
 \quad \quad \kvd{match}~\ident{Map.find}~\str{temp}~\ident{main}~\kvd{with} \\
 \quad \quad |~\ident{Number}(\ident{num})\rightarrow \ident{printfn}~\str{Lovely \%f!}~\ident{num} \\
 \quad \quad |~\_\rightarrow \ident{failwith}~\str{Incorrect format} \\
 \quad |~\_\rightarrow \ident{failwith}~\str{Incorrect format} \\
 |~\_\rightarrow \ident{failwith}~\str{Incorrect format}
\end{array}
\end{equation*}
%
The code assumes that the response has a particular shape described in the documentation. The
root node must be a record with a \strf{main} field, which has to be another record containing
a numerical \strf{temp} field representing the current temperature. When the shape is different,
the code fails. While not immediately unsound, the code is prone to errors if strings are
misspelled or incorrect shape assumed.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.80,trim=5mm 5mm 5mm 5mm,clip]{hierarchy.pdf} % left bottom right top
\end{center}
\vspace{-0.5em}
\caption{Important aspects of the preferred shape relation}
\label{fig:subtyping-diagram}
\end{figure}

@@

Such $e'$ is obtained by transforming sub-expressions of $e$ using one of the following
translation rules:
%
\begin{enumerate}
\item
$C[e]$ to $C[\kvd{match}~e~\kvd{with}~\ident{Some}(v) \rightarrow v~|~\ident{None} \rightarrow \ident{exn}]$
\item
$C[e]$ to $C[e.\ident{M}]$ where $\ident{M} = \tytagof(\sigma)$ for some $\sigma$
\item
$C[e]$ to $C[\ident{int}(e)]$
\end{enumerate}
\end{remark}

@@

\begin{itemize}
\item[--] Primitive shapes can become nullable (1), \ident{int} can become \ident{float} (3)
  or become a part of a labelled top type (2).
\item[--] Record shape fields can change shape (recursively) and record may become a part
  of a labelled top type (2).
\item[--] For list and nullable shapes, the shape of the value may change (we apply the
  transformations recursively).
\item[--] For the $\kvd{any}$ shape, the original code will continue to work (none of the labels is ever removed).
\end{itemize}
\vspace{-1.5em}
\end{proof}
%
\noindent
Intuitively, the first transformation is needed when the new sample makes a type optional.
This happens when it contains a \kvd{null} value or a record that does not contain a field
that all previous samples have. The second transformation is needed when a shape $\sigma$
becomes $\kvd{any}\langle\sigma, \ldots\rangle$ and the third one is needed when $\ident{int}$
becomes \ident{float}.
